---
title: "ANA 515 Assignment 2"
author: "Yatin Pawar"
date: "11/12/2021"
output: html_document
---

## Description of dataset

This dataset of “Airline safety” from Five Thirty Eight’s GitHub data repository. It is trying to answer the question, “should travelers avoid flying airlines that have had crashes in the past?” Accidents happen either due to human errors or due to natural events. Is it true that past data of accidents make Airline vulnerable? An effort made to study data from Aviation Safety Network’s database of 30 years for 56 airlines. The 30-year period down into two halves: first from 1985 to 1999, and then from 2000 to 2014. The comparison to see if data shows any improvements or degradation of safety measures. If we identify a correlation, that will imply that crash risk is persistent — predictable to some extent based on the airline. 

The data was collected in CSV file. The column names are Name of airline, Number of available seat kilometers per week, Incidents from 1985 to 1999, Fatal accidents from 1985 to 1999, Fatalities from 1985 to 1999, Incidents from 2000 to 2014, Fatal accidents from 2000 to 2014, and Fatalities from 2000 to 2014. Except Airline name, rest of the data is numeric. 

```{r setup, include=TRUE}
#Next line url is to get the data from GitHub to R Studio wd.
#using read.csv to read data from csv file
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv"

airline_safety <- read.csv(url)

#Need to install below libraries to run the functions within them.
library(tidyverse)
library(knitr)
library(bslib)

#Renaming better column names
names(airline_safety)[3]<-"Incidents_1985-99"
names(airline_safety)[4]<-"Fatal_Accidents_1985-99"
names(airline_safety)[5]<-"Fatalities_1985-99"
names(airline_safety)[6]<-"Incidents_2000-14"
names(airline_safety)[7]<-"Fatal_Accidents_2000-14"
names(airline_safety)[8]<-"Fatalities_2000-14"

```


```{r, echo=TRUE}
#This next chunk is inline code. Inline code puts the text with the output of the function in the document.
```

This dataframe has `r nrow(airline_safety)` rows and `r ncol(airline_safety)` columns. The names of the columns and a brief description of each are in the table below:

# Coulmn Names and Description


```{r airline_safety, echo=TRUE}
library(knitr)
col_desc_airline_safety<-data.frame(
Names = c("Airline","avail_seat_km_per_week","Incidents_1985-99","Fatal_Accidents_1985-99","Fatalities_1985-99","Incidents_2000-14","Fatal_Accidents_2000-14","Fatalities_2000-14"),
  Description = c(
  "Airline (asterisk indicates that regional subsidiaries are included)",
  "Available seat kilometers flown every week", 
  "Total number of incidents, 1985–1999",
  "Total number of fatal accidents, 1985–1999","Total number of fatalities, 1985–1999", "Total number of incidents, 2000–2014", "Total number of fatal accidents, 2000–2014", "Total number of fatalities, 2000–2014")
)

knitr::kable(head(col_desc_airline_safety[, 1:2]), "simple")

```


```{r, include=TRUE}
#Create a new dataset with the name 'safetydata2' from the dataset ‘airline_safety’ that shows columns   "Incidents_2000-14","Fatal_Accidents_2000-14","Fatalities_2000-14"
library(dplyr)
safetydata2 <- airline_safety %>%
  select("Incidents_2000-14", "Fatal_Accidents_2000-14","Fatalities_2000-14")

```

# Statistical summary of data from year 2000 to 2014 as below:
```{r, echo=TRUE}
#summary of the dataset safetydata2 
sumdata<-summary(safetydata2)

print (sumdata) 

```

